\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2019-04-18 archive.tex
scientific laws
added caption and latest deep program learning figures

\begin{figure}
  \includegraphics[width = 1.2\textwidth]{figures/DeepLogo_4_11}
  \includegraphics[width = \textwidth]{figures/dpl/list}
  \caption{}\label{exampleDSL}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tabular}{lll}\toprule
    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
    % physical laws programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Center-of-mass:\\
        $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Integrating position over time:\\
        $x_0$, $\dot{x}$, $\ddot{x}$, $t\mapsto \frac{1}{2}\ddot{x}t^2 + \dot{x}t + x_0$\\

        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

@@

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\popp{Library}}&
    % physical laws library
    \begin{tabular}{l}
      \begin{tabular}{l}
        \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
        \hspace{\helpSize}($f_0$: \emph{L-2 norm})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
        \hspace{\helpSize}($f_1$: \emph{Sum vector components})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
        \hspace{\helpSize}($f_2$: \emph{Dot product})
      \end{tabular}\\
      %(lambda (fold $0 (map (lambda (*. $0 0.)) (car $0)) (lambda (lambda (zip $0 $1 (lambda (lambda (+. $0 $1)))))))) 
      \begin{tabular}{l}
        \greenCode{$f_3($vs$)\,=\,$(fold vs (map ($\lambda$ (x) (* x 0))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(fold vs (}}\greenCode{(car vs))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($x$)\,=\,$(/ x (+ 1 1))}\\
        \hspace{\helpSize}($f_4$: \emph{Divide by two})
      \end{tabular}\\
      %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
      \begin{tabular}{l}
        \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
        \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($v$)\,=\,$(power ($f_0$ v) ($f_4$ 1))}\\
        \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}\\
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_4$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
        \hspace{\helpSize}($f_6$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
        \hspace{\helpSize}($f_7$: \emph{index})
      \end{tabular}

@@

>>>>

\begin{figure}
  \includegraphics[width = \textwidth]{figures/DeepLogo_4_15}
  \includegraphics[width = \textwidth]{figures/DeepList_4_17}
  \caption{Model solves tasks (left; top, task is to draw an image. bottom, task is input/output mapping) by writing programs (right column) where these programs invoke explicit declarative knowledge in the form of library routines (middle) jointly learned w/ the solutions. Library code can call each other (arrows), forming a layered network of domain-specific concepts. Fantasies (below library) are tasks sampled by randomly combining code from library. Fantasies are used to train a neural net to help search for programs.}\label{exampleDSL}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{tabular}{lll}\toprule
    &\multicolumn{1}{c}{Physical Laws}&\multicolumn{1}{c}{Recursive Functions}\\\midrule
    \normalsize \rotatebox[origin=c]{90}{\pop{Programs} \& Tasks}&
    % physical laws programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Kinetic energy:\\
        $m$, $\vec{v}\mapsto \frac{1}{2}m|\vec{v}|^2$\\
        \code{1, [0 2 -1]$\to$2.5}\\
        \code{2, [2 1 3]$\to$14}\\
        \blueCode{$f(m$, $\vec{v}) = $($f_4$ (* $m$ ($f_0$ $\vec{v}$)))}

        %% Center-of-mass:\\
        %% $m_1$, $\vec{r}_1$, $m_2$, $\vec{r}_2\mapsto \frac{m_1}{m_1 + m_2}\vec{r}_1 + \frac{m_2}{m_1 + m_2}\vec{r}_2$\\
        %% \code{2, [3 0 1], 4, [0 6 0]}$\to$\code{[1 4 0.333]}\\
        %% \code{5, [7 1 6], 5, [9 2 2]}$\to$\code{[8 0.5 4]}        
      \end{tabular}\\\\
      \begin{tabular}{l}
        Coulomb's law:\\
        $q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2\mapsto \frac{q_1q_2}{|\vec{r}_1 - \vec{r}_2|_2^2} \times \frac{\vec{r}_1 - \vec{r}_2}{|\vec{r}_1 - \vec{r}_2|}$\\
        \code{4, [1 5 2], 2, [2 5 2]$\to$[-8 0 0]}\\
        \code{-1, [0 1 1], -1, [0 0 0]$\to$[0 0.5 0.5]}        \\
        \blueCode{$f(q_1$, $\vec{r}_1$, $q_2$, $\vec{r}_2) = $}\\
        \hspace{0.2cm}\blueCode{($f_7$ $q_1$ $q_2$ (zip $\vec{r}_1$ $\vec{r}_2$ ($\lambda$ (x y) (- x y))))}
      \end{tabular}\\\\
      \begin{tabular}{l}
        Integrating angle over time:\\
        $\theta_0$, $\omega_0$, $\alpha$, $t\mapsto \frac{1}{2}\alpha t^2 + \omega_0t + \theta_0$\\
        \code{1, 3, 0, 2$\to$7}\\
        \code{-5, 0, 10, 1$\to$0}\\
        \blueCode{$f(\theta_0$, $\omega_0$, $\alpha$, t$) = $($f_5$ $\omega_0$ $\theta_0$ $\alpha$ t)}
        \end{tabular}
      \end{tabular}
    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        Duplicate each list element:\\
        \code{[9 2]$\to $[9 9 2 2]}\\
        \code{[5 0 3]$\to $[5 5 0 0 3 3]}\\
        \code{[1 2 3 4]$\to $[1 1 2 2 3 3 4 4]}\\
        \blueCode{$f(l) = $($f_0$ ($\lambda$ (x a) (cons x (cons x a))) l nil)}        
      \end{tabular}\\\\

@@

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\popp{Library}}&
    % physical laws library
    \begin{tabular}{l}
      \begin{tabular}{l}
        \greenCode{$f_0($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ (* x x) y)) v)}\\
        \hspace{\helpSize}($f_0$: \emph{L-2 norm})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_1($v$)\,=\,$(fold 0 ($\lambda$ (x y) (+ x y)) v)}\\
        \hspace{\helpSize}($f_1$: \emph{Sum vector components})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($v,u$)\,=\,$($f_1$ (zip v u ($\lambda$ (x y) (* x y))))}\\
        \hspace{\helpSize}($f_2$: \emph{Dot product})
      \end{tabular}\\
      %(lambda (fold $0 (map (lambda (*. $0 0.)) (car $0)) (lambda (lambda (zip $0 $1 (lambda (lambda (+. $0 $1)))))))) 
      \begin{tabular}{l}
        \greenCode{$f_3($vs$)\,=\,$(fold vs (map ($\lambda$ (x) (* x 0))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(fold vs (}}\greenCode{(car vs))}\\
        \phantom{\greenCode{$f_3($vs$)\,=\,$(}}\greenCode{($\lambda$ (v u) (zip v u ($\lambda$ (x y) (+ x y)))))}\\
        \hspace{\helpSize}($f_3$: \emph{Sum (list of) vectors})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($x$)\,=\,$(/ x (+ 1 1))}\\
        \hspace{\helpSize}($f_4$: \emph{Divide by two})
      \end{tabular}\\
      %      %#(lambda (lambda (lambda (lambda (+. $2 (*. $0 (+. (*. (#(lambda (/. $0 (+. 1. 1.))) $0) $1) $3))))))) $2 $3 $1 $0)))))
      \begin{tabular}{l}
        \greenCode{$f_5($v,x,a,t$)\,=\,$(+ x (* t (+ (* ($f_4$ t) a) v)))}\\
        \hspace{\helpSize}($f_5$: \emph{Integrate given first and second derivatives})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($v$)\,=\,$(power ($f_0$ v) ($f_4$ 1))}\\
        \hspace{\helpSize}($f_6$: \emph{n-dimensional vector length})
      \end{tabular}\\
      %(lambda (lambda (lambda (map (lambda (*. (/. (*. (/. $0 (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $1)) $3) (#(lambda (power (#(lambda (fold $0 0. (lambda (lambda (+. (*. $1 $1) $0))))) $0) (#(lambda (/. $0 (+. 1. 1.))) 1.))) $1)) $2)) $0)
    %(lambda (lambda (lambda (map ($\lambda$ (x) (* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)
    \begin{tabular}{l}
      \greenCode{$f_7($a,b,v$)\,=\,$(map ($\lambda$ (x)}\\
      \greenCode{\phantom{$f_7($a,b}(* (/ (* (/ x ($f_0$ v)) a) ($f_6$ v)) b)) v)}\\
      \hspace{\helpSize}($f_7$: Inverse square law)
      \end{tabular}
    \end{tabular}

    &
    % recursive functions programs and tasks
    \begin{tabular}{l}
      \begin{tabular}{l}
        \popp{\code{$f_0($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
        \phantom{\code{$f_0($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_0$ (cdr l))))}}\\
          \hspace{\helpSize}($f_0$: \emph{fold})
      \end{tabular}\\
      \begin{tabular}{l}
        \popp{\code{$f_1($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
        \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_1$ (n x))))}}\\
          \hspace{\helpSize}($f_1$: \emph{unfold})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_2($f$,$l$)\,=\,$($f_0$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_2$: \emph{map})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_3($f$,$p$,$n$)\,=\,$($f_1$ p f (+ 1) n)}\\
        \hspace{\helpSize}($f_3$: \emph{count upward until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_4($l$)\,=\,$($f_0$ ($\lambda$ (a x) (+ 1 a)) l 0)}\\
        \hspace{\helpSize}($f_4$: \emph{length})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_5($p$)\,=\,$($f_3$ ($\lambda$ (x) x) p 0)}\\
        \hspace{\helpSize}($f_5$: \emph{count upward from 0 until predicate holds})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_6($n$)\,=\,$($f_5$ (eq? n))}\\
        \hspace{\helpSize}($f_6$: \emph{range})
      \end{tabular}\\
      \begin{tabular}{l}
        \greenCode{$f_7($l$,$n$)\,=\,$(car ($f_0$ ($\lambda$ (a x) (cdr a)) ($f_6$ n) l))}\\
        \hspace{\helpSize}($f_7$: \emph{index})
      \end{tabular}

@@

Our model alternatingly searches for solutions to tasks and then improves its search strategy;
within program induction, these principles were first elucidated by the Exploration-Compression (EC) algorithm~\cite{Dechter:2013:BLV:2540128.2540316}.
